Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-02-07 parallel-tractability-J-v9/conclusions.tex
overall revision to make some presentations consistent

In this paper, we studied the problem of parallel tractability of
materialization on the datalog rewritable ontologies.
To identify the parallelly tractable classes,
we proposed two NC algorithms, Algorithm~$\mathsf{A}_{bsc}^\psi$
and Algorithm~$\mathsf{A}_{opt}^\psi$, that perform materialization on
datalog rewritable ontology languages.
Based on these algorithms, we identified the corresponding
parallelly tractable datalog program (\texttt{PTD}) classes such that materialization
on the datalog programs in these classes is in the complexity class NC.
We further studied two specific ontology languages, DL-Lite and DHL (including one of its extension).
We showed that any ontology expressed in DL-Lite$_{core}$ or DL-Lite$_{\mathcal{R}}$ is parallelly tractable.
For DHL and DHL($\circ$), we proposed two restrictions such that materialization is parallelly tractable.

We verified the usefulness of our theoretical techniques in two ways. On the one hand,
we analyzed different kinds of datasets,
including well-known benchmarks, real-world ontologies and a famous dataset YAGO.
Our analysis showed that many real ontologies belong to
the parallelly tractable class $\mathcal{D}_{\textit{\text{dhl}}}$
or $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$. The
developers and users can also refer to $\mathcal{D}_{\textit{\text{dhl}}}$
and $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$
to create large-scale ontologies for which parallel tractability
is theoretically guaranteed. On the other hand,
we used an optimization strategy based on SWD paths
to give a practical algorithm variant Algorithm~$\mathsf{A}_{prc}$, which can
also be restricted to an NC version, and can be implemented for practice.
We implemented a system based on Algorithm~$\mathsf{A}_{prc}$ and compared it to
the state-of-the-art reasoner RDFox.
The experimental results showed that the optimization proposed in this paper results
in a better performance on parallelly tractable ontologies compared with RDFox.

In the future work, we will extend our work in two lines.
One line is a further study of the parallel tractability of
other expressive ontology languages, in addition to
datalog rewritable ones. Since different expressive OWL languages
have different syntaxes and higher reasoning complexities than \texttt{PTime}-complete, we need to
explore more restrictions that are practical and make materialization
parallelly tractable. Another line of the future work is to
study how to further apply the
results in practice.
One idea is to apply the technique of SWD paths to enhance other reasoning-based tasks,
like ontology classification, ontology debugging and query answering.

>>>>

In this paper, we studied the problem of parallel tractability of
materialization on the datalog rewritable ontologies.
To identify the parallel tractability classes,
we proposed two \texttt{NC} algorithms, Algorithm~$\mathsf{A}_{bsc}^\psi$
and Algorithm~$\mathsf{A}_{opt}^\psi$, that perform materialization on
datalog rewritable ontology languages.
Based on these algorithms, we identified the corresponding
parallel tractability classes such that materialization
on the datalog programs in these classes is in the NC complexity.
We further studied two specific ontology languages, DL-Lite and DHL (including one of its extension).
We showed that any ontology expressed in DL-Lite$_{core}$ or DL-Lite$_{\mathcal{R}}$ is tractable in parallel.
For DHL and DHL($\circ$), we proposed two restrictions such that materialization is tractable in parallel.

We analyzed the benchmark LUBM and the real-world dataset YAGO
and gave the cases where these ontologies belong to parallel tractability classes.
On the other hand, we used an optimization strategy based on SWD paths
to give a practical algorithm variant Algorithm~$\mathsf{A}_{prc}$, which can
also be restricted to an \texttt{NC} version.
We implemented a system based on Algorithm~$\mathsf{A}_{prc}$ and compared it to
the state-of-the-art reasoner RDFox.
The experimental results showed that our system is comparable to RDFox when handling
the LUBM and the YAGO ontologies.

In the future work, we will extend our work in two lines.
One line is a further study of the parallel tractability of
other expressive ontology languages, in addition to
datalog rewritable ones. Since different expressive OWL languages
have different syntaxes and higher reasoning complexities than \texttt{PTime}-complete, we need to
explore more restrictions that are practical and make materialization
tractable in parallel. Another line of the future work is to
study how to further apply the
results in practice.
One idea is to apply the technique of SWD paths to enhance other reasoning-based tasks,
like ontology classification, ontology debugging and query answering.

### 2018-02-07 parallel-tractability-J-v9/evaluation.tex
overall revision to make some presentations consistent

\section{Evaluation}
\label{sec:evaluation}

In the first part of this section, we analyze different kinds of ontologies and datasets
and investigate the usability of the results proposed above.
In the second part, we evaluate the implementation of Algorithm~$\mathsf{A}_{\text{prc}}$
and compare it to the reasoning system RDFox.

@@

In this part, we analyze two popular datasets, LUBM and YAGO, and show the relations
between these two datasets and the theoretical results regarding parallel tractability.

\textbf{LUBM}. In the Semantic Web community, LUBM
(The Lehigh University Benchmark) is proposed to
facilitate the evaluation of ontology-based systems
in a standard and systematic way.
In the latest version of LUBM,\footnote{http://swat.cse.lehigh.edu/projects/lubm/}
the core ontology contains 48 classes and 32 properties, used to describe the departments and the staff of
universities. By setting a different number of universities for an ontology-generator, users can get datasets of any size based on the core ontology.
%
The statements about properties in the LUBM core ontology, such as inverse property statements,
can be rewritten into the datalog rules of the form (R1), (R2) and (R3) in Table~\ref{tab:dhl}.
Most of the statements about classes can be rewritten into the datalog rules of the form (T1) and (T3)
in Table~\ref{tab:dhl}. Five axioms have, however, the form $A\sqsubseteq\exists R.B$,
which requires existentially quantified variables in the rule head when rewriting
the axiom into a logic rule: $A(x)\rightarrow\exists y(R(x,y)\wedge B(y))$ ($\tau$),
where a free variable $y$ is introduced. This kind of axiom is a general case of (T4) in Table~\ref{tab:dhl}
for which $A$ is actually replaced by the top concept $\top$.
Similarly to how we handle (T4), we can also eliminate the free variable $y$
in rule $\tau$ via Skolemization, i.e., by replacing the variable $y$ with a new constant $o_R^{A,B}$.
In this way, rule $\tau$ can be rewritten into $A(x)\rightarrow R(x,o_R^{A,B})\wedge B(o_R^{A,B})$.
If we only focus on the materialization task, the rewriting approach via Skolemization guarantees the
completeness and correctness \cite{GrauHKKMMW13}.
On the other hand, rule $\tau$ is not considered when using OWL RL reasoners to handle LUBM \cite{UrbaniKMHB12,WeaverH09}.
In summary, if the rewriting approach is used for the above kind of rule,
the materialization of a LUBM dataset can be handled by
Algorithm~$\mathsf{A}_{\text{prc}}$.

\textbf{YAGO}. The knowledge base YAGO\footnote{http://www.mpi-inf.mpg.de/home/}
is constructed from Wikipedia and WordNet. The latest version
YAGO3 \cite{MahdisoltaniBS15} has more than 10 million entities
(e.g., persons, organizations, cities, etc.)
and contains more than 120 million facts about these entities.

In order to balance the expressiveness and computing efficiency,
a YAGO-style language, called the \emph{YAGO model}, is proposed based on
a slight extension of RDFS \cite{SuchanekKW08}. The YAGO model defines
a set of properties: \texttt{domain, range, subClassOf, subRelationOf} and \texttt{type},
and a set of classes: \texttt{entity, class, relation} and \texttt{acyclicTransitiveRelation}.
The facts in the YAGO model are stated by triples, e.g., $(r_1,
\texttt{subRelationOf}, r_2)$,
which are similar to the RDFS statements.
A group of rules for reasoning over YAGO ontologies
is specified as follows \cite{SuchanekKW08}:
\begin{enumerate}[leftmargin=8ex,label=(\arabic*),ref=\arabic*]
  \item $(r,\texttt{domain}, c), (x, r, y) \rightarrow (x,
    \texttt{type}, c)$\label{yago:r1}
  \item $(r,\texttt{range}, c), (x, r, y) \rightarrow (y,
    \texttt{type}, c)$\label{yago:r2}
  \item $(c_1, \texttt{subClassOf}, c_2), (x, \texttt{type}, c_1)
    \rightarrow (x, \texttt{type}, c_2)$\label{yago:r3}
  \item $(r_1, \texttt{subRelationOf}, r_2), (x, r_1, y) \rightarrow
    (x, r_2, y)$\label{yago:r4}
  \item $(r, \texttt{type}, \texttt{acyclicTransitiveRelation}), (x,
    r, y), (y, r, z) \rightarrow (x, r, z)$\label{yago:r5}
\end{enumerate}

@@

We run ParallelDHL on LUBM and YAGO to further analyze these
two kinds of ontologies. We also use the reasoning system, RDFox \cite{MotikNPHO14},
for comparison. We use five LUBM ontologies, lubm-50, lubm-100, lubm-150, lubm-200
and lubm-250, where lubm-50 contains the descriptions of 50 universities (the other ontologies
are explained similarly). For YAGO, we use its core version, denoted by yago-core.\footnote{
The yago-core ontology is available at https://www.mpi-inf.mpg.de/departments/databases-and-information-systems/research/yago-naga/yago/downloads/.}
The yago-core ontology is a subset of the full YAGO dataset. It does not include the datatype statements, the links
between different data sources, the degrees of confidence and other kinds of
annotations that we do not consider in this work. The statistics of the above ontologies is given in Table~\ref{tab:onto}.
The running platform for this experiment is a server with a 50 Gigabyte RAM and 8 physical cores, in each core
three logic threads can be allocated.

We run the LUBM ontologies and the yago-core ontology on RDFox and ParallelDHL with different numbers of
threads (respectively 1,4,8,16 and 24).
The reasoning times are collected in Table~\ref{tab:result}. We also give Figure~\ref{fig:reasoningtime} to
clearly compare the two systems.
In Figure~\ref{fig:reasoningtime}(left), the abscissa records the five LUBM ontologies,
the ordinate records the reasoning times with 24 threads being allocated;
In Figure~\ref{fig:reasoningtime}(right), the abscissa records the reciprocal of the number of threads, and
the ordinate records the reasoning times over the yago-core ontology.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1\textwidth]{fig-reasoningtime.eps}
\caption{(left) the reasoning times over LUBM ontologies; (right) the reasoning times over yago-core.}
\label{fig:reasoningtime}
\end{center}
\end{figure}

@@

From the above results, we can find that the graph depths are far less than
the sizes of the input ontologies. Since the graph depth determines the computing
time of Algorithm~$\mathsf{A}_{bsc}$ and Algorithm~$\mathsf{A}_{opt}$ as discussed
at the beginning of this part,
the test ontologies should be materialized efficiently using these two algorithms.
The obvious difference of the reasoning times of the test ontologies is due to the limited memory space and the fixed number
of processors and the inapplicability of the parallel assumption given in Section~\ref{sec:alg-bsc}.
This results in that the computing
time of ParallelDHL does not depend only on the number of iterations of \ref{alg4:updateG}
of Algorithm~$\mathsf{A}_{prc}$ as analyzed in Section~\ref{sec:practicalAlg}.
It depends on several factors: the balancing of workload, the delay of some processors,
the I/O blocking, etc. Since these issues are not our focus, we do not analyze them in this work.
On the other hand, the results of the graph depths
indicate that we can improve the performance of the materialization task by using more
memory space and more processors.

>>>>

\section{Evaluation and Analysis}
\label{sec:evaluation}

In the first part of this section, we analyze the well-known benchmark, LUBM, 
and the real-world dataset, YAGO, and
show the cases where these ontologies belong to the parallel tractability classes.
In the second part, we evaluate the implementation of Algorithm~$\mathsf{A}_{\text{prc}}$
and compare it to the reasoning system RDFox.

@@

\textbf{LUBM}. In the Semantic Web community, LUBM
(The Lehigh University Benchmark) is proposed to
facilitate the evaluation of ontology-based systems
in a standard and systematic way.
In the latest version of LUBM,\footnote{http://swat.cse.lehigh.edu/projects/lubm/}
the core ontology contains 48 classes and 32 properties, used to describe the departments and the staff of
universities. By setting a different number of universities for an ontology-generator, users can get datasets of any size based on the core ontology.
%
The statements about properties in the LUBM core ontology, such as inverse property statements,
can be rewritten into the datalog rules of the form (R1), (R2) and (R3) in Table~\ref{tab:dhl}.
Most of the statements about classes can be rewritten into the datalog rules of the form (T1) and (T3)
in Table~\ref{tab:dhl}. Five axioms have, however, the form $A\sqsubseteq\exists R.B$,
which requires existentially quantified variables in the rule head when rewriting
the axiom into a logic rule: $A(x)\rightarrow\exists y(R(x,y)\wedge B(y))$ ($\tau$),
where a free variable $y$ is introduced. This kind of axiom is a general case of (T4) in Table~\ref{tab:dhl}
for which $A$ is actually replaced by the top concept $\top$.
Similarly to how we handle (T4), we can also eliminate the free variable $y$
in rule $\tau$ via Skolemization, i.e., by replacing the variable $y$ with a new constant $o_R^{A,B}$.
In this way, rule $\tau$ can be rewritten into $A(x)\rightarrow R(x,o_R^{A,B})\wedge B(o_R^{A,B})$.
If we only focus on the materialization task, the rewriting approach via Skolemization guarantees the
completeness and correctness \cite{GrauHKKMMW13}.
On the other hand, rule $\tau$ is not considered when using OWL RL reasoners to handle LUBM \cite{UrbaniKMHB12,WeaverH09}.
In summary, if the rewriting approach is used for the above kind of rule,
the materialization of LUBM datasets is tractable in parallel and can be handled by
Algorithm~$\mathsf{A}_{\text{prc}}$.

\textbf{YAGO}. The knowledge base YAGO\footnote{http://www.mpi-inf.mpg.de/home/}
is constructed from Wikipedia and WordNet. The latest version
YAGO3 \cite{MahdisoltaniBS15} has millions of facts.
In order to balance the expressiveness and computing efficiency,
a YAGO-style language, called the \emph{YAGO model}, is proposed based on
a slight extension of RDFS \cite{SuchanekKW08}. The YAGO model defines
a set of properties: \texttt{domain, range, subClassOf, subRelationOf} and \texttt{type},
and a set of classes: \texttt{entity, class, relation} and \texttt{acyclicTransitiveRelation}.
The facts in the YAGO model are stated by triples, e.g., $(r_1,
\texttt{subRelationOf}, r_2)$,
which are similar to the RDFS statements.
A group of rules for reasoning over YAGO ontologies
is specified as follows \cite{SuchanekKW08}:
\begin{enumerate}[leftmargin=8ex,label=(\arabic*),ref=\arabic*]
  \item $(r,\texttt{domain}, c), (x, r, y) \rightarrow (x,
    \texttt{type}, c)$\label{yago:r1}
  \item $(r,\texttt{range}, c), (x, r, y) \rightarrow (y,
    \texttt{type}, c)$\label{yago:r2}
  \item $(c_1, \texttt{subClassOf}, c_2), (x, \texttt{type}, c_1)
    \rightarrow (x, \texttt{type}, c_2)$\label{yago:r3}
  \item $(r_1, \texttt{subRelationOf}, r_2), (x, r_1, y) \rightarrow
    (x, r_2, y)$\label{yago:r4}
  \item $(r, \texttt{type}, \texttt{acyclicTransitiveRelation}), (x,
    r, y), (y, r, z) \rightarrow (x, r, z)$\label{yago:r5}
\end{enumerate}

@@

We run ParallelDHL on the LUBM and the YAGO ontologies to further analyze these
two kinds of ontologies. We also use the reasoning system, RDFox \cite{MotikNPHO14},
for comparison. We use five LUBM ontologies, lubm-50, lubm-100, lubm-150, lubm-200
and lubm-250, where lubm-50 contains the descriptions of 50 universities (the other ontologies
are explained similarly). For YAGO, we use its core version, denoted by yago-core.\footnote{
The yago-core ontology is available at https://www.mpi-inf.mpg.de/departments/databases-and-information-systems/research/yago-naga/yago/downloads/.}
The yago-core ontology is a subset of the full YAGO dataset without imports. It does not include the datatype statements, the links
between different data sources, the degrees of confidence and other kinds of
annotations that we do not consider in this work. The statistics of the above ontologies is given in Table~\ref{tab:onto}.
The running platform for this experiment is a server with a 50 Gigabyte RAM and 8 physical cores, in each core
three logic threads can be allocated.

We run the LUBM ontologies and the yago-core ontology on RDFox and ParallelDHL with different numbers of
threads (respectively 1,4,8,16 and 24).
The reasoning times are collected in Table~\ref{tab:result}. We also give Figure~\ref{fig:reasoningtime} to
clearly compare the two systems.
In Figure~\ref{fig:reasoningtime}(left), the abscissa records the five LUBM ontologies,
the ordinate records the reasoning times with 24 threads being allocated;
In Figure~\ref{fig:reasoningtime}(right), the abscissa records the reciprocals of the numbers of threads, and
the ordinate records the reasoning times over the yago-core ontology.

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1\textwidth]{fig-reasoningtime.eps}
\caption{(left) the reasoning times over the LUBM ontologies; (right) the reasoning times over yago-core.}
\label{fig:reasoningtime}
\end{center}
\end{figure}

@@

From the above results, we can find that the graph depths are far less than
the sizes of the input ontologies. Since the graph depth determines the computing
time of Algorithm~$\mathsf{A}_{bsc}$ and Algorithm~$\mathsf{A}_{opt}$ as discussed
at the beginning of this part,
the test ontologies should be materialized efficiently using these two algorithms.
The obvious difference of the reasoning times of the test ontologies is due to the limited memory space and the fixed number
of processors and the inapplicability of the parallel assumption given in Section~\ref{sec:alg-bsc}.
This results in that the computing
time of ParallelDHL does not depend only on the number of iterations of \ref{alg4:updateG}
of Algorithm~$\mathsf{A}_{prc}$ as analyzed in Section~\ref{sec:practicalAlg}.
It depends on several factors: the balancing of workload, the delay of some processors,
the I/O blocking, etc. Since these issues are not our focus, we do not analyze them in this work.
On the other hand, the results of the graph depths
indicate that we can further improve the performance of the materialization task by using more
memory space and more processors.

### 2018-02-07 parallel-tractability-J-v9/introduction.tex
overall revision to make some presentations consistent

The effectiveness of parallel reasoning for the above mentioned techniques is empirically verified on different test datasets.
However, materialization is not tractable in parallel for most of the popular ontology languages with \texttt{PTime}-complete or higher complexity \cite{Raymond95}. In particular, this is the case for RDFS and datalog rewritable ontology languages and, therefore, the efficiency of reasoning may not be
improved using a parallel implementation. A possible reason for the apparent high performance of parallel
reasoning is that the utilized test datasets do not fall into the worst cases in terms of computational complexity. Also some well-known, large-scale ontologies such as YAGO show good performance for parallel
reasoning \cite{KolovskiWE10}, although they are expressed in ontology languages that are, in theory, not tractable in parallel.
On the other hand, there are ontologies for which materialization performance does not improve by parallelism and our experiments in Section~\ref{sec:evaluation} confirm this.
The current theoretical work of parallel ontology reasoning can hardly explain this.
While one can try out different parallel implementations to see whether an ontology can be handled by (one of) them efficiently,
we study the problem of parallel tractability in theory and identify properties that make an ontology tractable in parallel.
These properties can further be used to optimize parallel algorithms and to guide ontology engineers in creating ontologies
for which parallel tractability can be guaranteed theoretically.

@@

In the practical part of this work, we study specific datalog rewritable ontology languages.
We first focus on the ontology language DL-Lite \cite{CalvaneseGLLR07} to clarify how to apply the above \texttt{NC} algorithms
to study parallel tractability of ontology materialization.
We show that DL-Lite$_{\text{core}}$ and DL-Lite$_\mathcal{R}$ are tractable in parallel.
We then study the ontology language Description Horn Logic (DHL) \cite{GrosofHVD03}, which is the intersection of datalog and OWL
in terms of expressivity. We give a case of a DHL ontology where materialization can hardly be parallelized.
Based on the analysis of this case, we propose to restrict the usage of DHL such that materialization over the
restricted ontologies can be handled by the proposed \texttt{NC} algorithms.
We further extend the results to an extension of DHL that also allows complex role inclusion axioms.
Finally, we analyze well-known benchmarks and real-world datasets and show that many real ontologies following the
proposed restrictions belong to the parallel tractability classes.
We implement a system based on an optimized \texttt{NC} algorithm for DHL materialization, which we compare to the state-of-the-art reasoner RDFox.
The experimental results show that the optimizations proposed in this paper result in a better performance on ontologies that are 
tractable in parallel compared with RDFox.

The remainder of the paper is organized as follows. In Section~\ref{sec:background}, we introduce some basic notions.
We then give two \texttt{NC} algorithms for ontology materialization in Section~\ref{sec:ptclass}.
We study parallel tractability of materialization in DL-Lite, DHL and an extension of DHL in Section~\ref{sec:ptonto}, respectively.
In Section~\ref{sec:practicalAlg},
we discuss how to optimize and implement the given \texttt{NC} algorithms.
We analyze real-world datasets and evaluate our implementation in Section~\ref{sec:evaluation}.
Finally, we discuss related work in Section~\ref{sec:related} and conclude in Section~\ref{sec:conclusion}. Detailed proofs of the theorems and lemmas in this paper are given in the appendix.

>>>>

The effectiveness of parallel reasoning for the above mentioned techniques is empirically verified on different test datasets.
However, materialization is not tractable in parallel for most of the popular ontology languages with \texttt{PTime}-complete or higher complexity \cite{Raymond95}. In particular, this is the case for RDFS and datalog rewritable ontology languages and, therefore, the efficiency of reasoning may not be
improved using a parallel implementation in theory. A possible reason for the apparent high performance of parallel
reasoning is that the utilized test datasets do not fall into the worst cases in terms of computational complexity. Also some well-known, large-scale ontologies such as YAGO show good performance for parallel
reasoning \cite{KolovskiWE10}, although they are expressed in ontology languages that are, in theory, not tractable in parallel.
The current theoretical work of parallel ontology reasoning can hardly explain this.
While one can try out different parallel implementations to see whether an ontology can be handled by (one of) them efficiently,
we study the problem of parallel tractability in theory and identify properties that make an ontology tractable in parallel.
These properties can further be used to optimize parallel algorithms and to guide ontology engineers in creating ontologies
for which parallel tractability can be guaranteed theoretically.

@@

In the practical part of this work, we study specific datalog rewritable ontology languages.
We first focus on the ontology language DL-Lite \cite{CalvaneseGLLR07} to clarify how to apply the above \texttt{NC} algorithms
to study parallel tractability of ontology materialization.
We show that DL-Lite$_{\text{core}}$ and DL-Lite$_\mathcal{R}$ are tractable in parallel.
We then study the ontology language Description Horn Logic (DHL) \cite{GrosofHVD03}, which is the intersection of datalog and OWL
in terms of expressivity. We give a case of a DHL ontology where materialization can hardly be parallelized.
Based on the analysis of this case, we propose to restrict the usage of DHL such that materialization over the
restricted ontologies can be handled by the proposed \texttt{NC} algorithms.
We further extend the results to an extension of DHL that also allows complex role inclusion axioms.
Finally, we analyze the well-known benchmark, LUBM, and the real-world dataset, YAGO, and 
show that these ontologies following the
proposed restrictions belong to the parallel tractability classes.
We implement a system based on an optimized \texttt{NC} algorithm for DHL materialization
and show that this system is comparable to the state-of-the-art reasoner RDFox
when handling the LUBM and the YAGO ontologies.

The remainder of the paper is organized as follows. In Section~\ref{sec:background}, we introduce some basic notions.
We then give two \texttt{NC} algorithms for ontology materialization in Section~\ref{sec:ptclass}.
We study parallel tractability of materialization in DL-Lite, DHL and an extension of DHL in Section~\ref{sec:ptonto}, respectively.
In Section~\ref{sec:practicalAlg},
we discuss how to optimize and implement the given \texttt{NC} algorithms.
We analyze the LUBM and the YAGO ontologies and evaluate our implementation in Section~\ref{sec:evaluation}.
Finally, we discuss related work in Section~\ref{sec:related} and conclude in Section~\ref{sec:conclusion}. Detailed proofs of the theorems and lemmas in this paper are given in the appendix.

### 2018-02-07 parallel-tractability-J-v9/parallel-tractability-J.tex
worked on experimental section, fixed table layout and captions
added the experiments of lubm and yago
overall revision to make some presentations consistent

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}

\usepackage{amsmath}

\usepackage{CJK}
\usepackage{color}

@@

@@

\begin{abstract}
  Materialization is an important reasoning service for many
  ontology-based applications, but the rapid growth of semantic data
  poses the challenge to efficiently perform materialization on
  large-scale ontologies. Parallel materialization algorithms work
  well for some ontologies, which is not always explained by the
  theoretical results. For example, for the YAGO ontology
  parallelization works well, although the reasoning problem for the
  used ontology language is not in \texttt{NC}, i.e., the theoretical
  complexity class for \emph{parallel tractability}. This motivates us
  to study the problem of \emph{parallel tractability} of ontology
  materialization from a theoretical perspective. We focus on the
  datalog rewritable ontology languages DL-Lite and Description Horn
  Logic (DHL) and propose algorithms, called \texttt{NC} algorithms,
  to identify classes of ontologies for which materialization is
  tractable in parallel. To verify the practical usability of the
  above results, we analyze real-world datasets and show that for many
  ontologies expressed in DHL materialization is tractable in
  parallel. The implementation of our optimized parallel algorithm
  shows performance improvements over the highly optimized
  state-of-the-art reasoner RDFox on ontologies for which
  materialization is tractable in parallel.
\end{abstract}

@@

>>>>

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{CJK}
\usepackage{color}

@@

\usepackage{array}
\usepackage{subcaption}
\captionsetup[subfigure]{skip=-10pt}

@@

\begin{abstract}
  Materialization is an important reasoning service for many
  ontology-based applications, but the rapid growth of semantic data
  poses the challenge to efficiently perform materialization on
  large-scale ontologies. Parallel materialization algorithms work
  well for some ontologies, although the reasoning problem for the
  used ontology language is not in \texttt{NC}, i.e., the theoretical
  complexity class for \emph{parallel tractability}. This motivates us
  to study the problem of \emph{parallel tractability} of ontology
  materialization from a theoretical perspective. We focus on the
  datalog rewritable ontology languages DL-Lite and Description Horn
  Logic (DHL) and propose algorithms, called \texttt{NC} algorithms,
  to identify classes of ontologies for which materialization is
  tractable in parallel. To verify the practical usability of the
  above results, we analyze benchmark and real-world datasets, including the LUBM
  and the YAGO ontologies, and show that for many
  ontologies expressed in DHL materialization is tractable in
  parallel. The implementation of our optimized parallel algorithm
  shows comparable performance with compared to the highly optimized
  state-of-the-art reasoner RDFox.
\end{abstract}

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "parallel-tractability-J"
%%% End:

### 2018-02-07 parallel-tractability-J-v9/parallelTractabilityDatalog.tex
overall revision to make some presentations consistent

Our target is to find for which kinds of ontologies (not ontology languages)
materialization is tractable in parallel. We consider data complexity,
i.e., we consider classes of datalog programs which share a fixed the rule set.
The data complexity of the materialization problem in this case is
$\texttt{PTime}$-complete, which is considered to be inherently sequential in the worst
case \cite{Raymond95}.
% Since we assume that, for any class of datalog programs where the rule
% set of each datalog program is fixed, the materialization problem is, thus, in data complexity
% $\texttt{PTime}$-complete, which is considered to be inherently sequential in the worst
% case \cite{Raymond95}.
In other words, the materialization problem of
datalog programs cannot be solved in parallel poly-logarithmic time unless \texttt{P}=\texttt{NC}.
Thus, we say that \emph{materialization for a class of datalog programs is tractable in parallel
if there exists an algorithm that handles this class of datalog programs and runs in parallel
poly-logarithmic time. Such an algorithm is also called an \texttt{NC} algorithm.}
In this section, we identify such classes of datalog programs by studying different \texttt{NC} algorithms.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{fig-general.pdf}
\caption{A partial materialization graph}
\label{fig:general}
\end{center}
\end{figure}

@@

\begin{proof}[Proof sketch]
Suppose $P\in\mathcal{D}_{\mathsf{A}_{\text{bsc}}^{\psi}}$. According to Theorem~\ref{theorem:a1},
the depth of the materialization graph $\mathcal{G}$ constructed by $\mathsf{A}_{\text{bsc}}^{\psi}$ is upper-bounded by $\psi(|P|)$.
It is obvious that the number of nodes in each path of $\mathcal{G}$ is also upper-bounded by $\psi(|P|)$.
According to the optimization strategy applied in $\mathsf{A}_{\text{opt}}$, if $\mathcal{G}$ can be constructed
by $\mathsf{A}_{\text{opt}}$, the number of iterations of $\mathsf{A}_{\text{opt}}$ has to be upper-bounded by $\psi(|P|)$;
if $\mathcal{G}$ is not the materialization graph constructed
by $\mathsf{A}_{\text{opt}}$, then there has to exist another materialization graph $\mathcal{G}'$ constructed by $\mathsf{A}_{\text{opt}}$
and $\mathcal{G}'$ has a smaller depth compared to $\mathcal{G}$.
\end{proof}

>>>>

Our target is to find for which kinds of ontologies (not ontology languages)
materialization is tractable in parallel. We consider data complexity,
i.e., we consider classes of datalog programs which share a fixed rule set.
The data complexity of the materialization problem in this case is
$\texttt{PTime}$-complete, which is considered to be inherently sequential in the worst
case \cite{Raymond95}.
% Since we assume that, for any class of datalog programs where the rule
% set of each datalog program is fixed, the materialization problem is, thus, in data complexity
% $\texttt{PTime}$-complete, which is considered to be inherently sequential in the worst
% case \cite{Raymond95}.
In other words, the materialization problem of
datalog programs cannot be solved in parallel poly-logarithmic time unless \texttt{P}=\texttt{NC}.
Thus, we say that \emph{materialization for a class of datalog programs is tractable in parallel
if there exists an algorithm that handles this class of datalog programs and runs in parallel
poly-logarithmic time. Such an algorithm is also called an \texttt{NC} algorithm.}
In this section, we identify such classes of datalog programs by studying different \texttt{NC} algorithms.

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{fig-general.pdf}
\caption{A partial materialization graph}
\label{fig:general}
\end{center}
\end{figure}

@@

\begin{proof}[Proof sketch]
Suppose $P=\langle R,\textbf{I}\rangle\in\mathcal{D}_{\mathsf{A}_{\text{bsc}}^{\psi}}$. According to Theorem~\ref{theorem:a1},
the depth of the materialization graph $\mathcal{G}$ constructed by $\mathsf{A}_{\text{bsc}}^{\psi}$ is 
upper-bounded by $\psi(|\textbf{I}|)$.
It is obvious that the number of nodes in each path of $\mathcal{G}$ is also upper-bounded by $\psi(|\textbf{I}|)$.
According to the optimization strategy applied in $\mathsf{A}_{\text{opt}}$, if $\mathcal{G}$ can be constructed
by $\mathsf{A}_{\text{opt}}$, the number of iterations of $\mathsf{A}_{\text{opt}}$ has to be upper-bounded by $\psi(|\textbf{I}|)$;
if $\mathcal{G}$ is not the materialization graph constructed
by $\mathsf{A}_{\text{opt}}$, then there has to exist another materialization graph $\mathcal{G}'$ constructed by $\mathsf{A}_{\text{opt}}$
and $\mathcal{G}'$ has a smaller depth compared to $\mathcal{G}$.
\end{proof}
